---
title: "F2 maternal effect"
author: "Erik Parker"
date: "May 11, 2018"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, tidy = TRUE)
```

```{r, echo = FALSE}


library(ggpubr)
library(gridExtra)
library(tidyr)
library(alr4)
library(visreg)
library(plyr)
library(relaimpo)
library(ggsignif)
library(lme4)
library(faraway)
library(RLRsim)
library(ggfortify)


f2 <- read.table("f2validation.csv", sep = ",", header = TRUE)

f2.validate <- f2[complete.cases(f2[,8]),]

f2.validate <- separate(data = f2.validate, col = Mother.Individual, into = c("mother", "individual"), sep = "\\.")


theme_Publication <- function(base_size=14, base_family="helvetica") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}


```


```{r}

ggplot(f2.validate, aes(x = Maternal.size, y = Scope.size)) + geom_point() + geom_jitter() + geom_smooth(method = "lm") + labs(y = "Offspring size", x = "Maternal size")

```

> First, though there are unfortunately not many mothers in this analysis, we see that we can still recover the normal trend of increasing maternal size -> increasing offspring size as we expect. As we knew going in that we didn't have a lot of mothers alive for this part of the study, we measured individual brood ball sizes of the F2s because this is generally taken as the primary way that mothers invest in their offspring. Doing this, we were still able to control for maternal effects - even without a lot of variation in maternal size.

```{r}

ggplot(f2.validate, aes(x = Treatment.ID, y = BB.weight)) + geom_boxplot() + geom_signif(comparisons = list(c("gg","gs")), test = "wilcox.test", map_signif_level = FALSE, textsize = 7, vjust = +1.2) + labs(y = "BB weight", x = "Microbiota treatment")

ggplot(f2.validate, aes(x = Treatment.ID, y = Scope.size)) + geom_boxplot() + geom_signif(comparisons = list(c("gg","gs")), test = "wilcox.test", map_signif_level = FALSE, textsize = 7, vjust = +1.2) + labs(y = "Offspring size", x = "Microbiota treatment")

```

> Also, we do see a significant difference (from a Wilcoxon signed-rank test) in the brood ball sizes produced by the mothers who received different treatments early in development, and also a significant difference in offspring size of mothers from different treatment groups. This together suggests to me that there is likely an interaction between our proxy for a maternal effect, brood ball size, and treatment.


```{r}

m3 <- lm(Scope.size ~ Sex + Treatment.ID + BB.weight + Treatment.ID:BB.weight + Sex:Treatment.ID, data = f2.validate)

Anova(m3)

ggplot(f2.validate, aes(x = Treatment.ID, y = BB.weight)) + geom_boxplot()

ggplot(f2.validate, aes(x = Treatment.ID, y = Scope.size, fill = Sex)) + geom_boxplot()

```

> Through model selection, I find that the best model for the raw data is one which contains an interaction between brood ball size and treatment (as well as one between sex and treatment to a lesser extent). This is not really suprising given the plots shown above. However, when we check this model for outlying and influential points, there is reason to be concerned with the conclusions of this model.

```{r}

plot(cooks.distance(m3), pch="*", cex=2, main="Influential Obs by Cooks distance")
abline(h = 4*mean(cooks.distance(m3), na.rm=T), col="red")
text(x=1:length(cooks.distance(m3))+1, 
     y=cooks.distance(m3), labels=ifelse(cooks.distance(m3)>4*mean(cooks.distance(m3), na.rm=T),
                                         names(cooks.distance(m3)),""), col="red")

ggplot(f2.validate, aes(x = BB.weight, y = Scope.size, color = Treatment.ID, label = rownames(f2.validate))) + geom_point() + geom_text(aes(label=rownames(f2.validate)),hjust=0, vjust=0) + labs(y = "Offspring size", x = "BB weight")

ggplot(f2.validate, aes(x = BB.weight, y = Scope.size, color = Treatment.ID)) + geom_point() + labs(y = "Offspring size", x = "BB weight") + geom_smooth(method = "lm")

```

> First, we see that a few datapoints are highly infleuntial by cook's distance. The 3 most infleutial seem to be observations 79, 89, and 98. Then, from an analysis of the plot of offspring size vs. bb weight, we see why. Observations 98 and 79 are quite clearly outliers - though 89 doesn't seem too bad, it is much closer to the rest of the dsitribution. Furthermore it's not as clearcut but obserbation 4, which was not identified as highly inflential by cook's distance, could also be considered an outlier.  
> The worst points are definitely observations 98 and 79 though, it is clear to see that they are likely largely responsible for the significant interaction between bb size and treatment identified in the previous model. I thus decided to remove them and go through another round of model selection to test this theory.  
> The outcome of this was basically that depending on which of these potential outliers I decided to remove, the interaction between bb size and treatment was either found to be significant or not. If I remove observations 98 and 79 (the clearest outliers), the interaction is still "significant", but just barely. But if I additionally remove observation 4 which is *extremely* borderline, and I keep going back and forth on the merits of including it, the interaction between bb size and treatment is no longer significant, and the model is just reduced to sex and treatment (what is currently presented in the paper). What this suggests to me is that there is *very likely* a maternal effect here in reality, it just isn't captured convincingly in these data due to the small sample size.  
> All that said, I'm still not completely sure how comfortable I am leaving observation 4 in the data (and thus including the maternal effect), as it alone causes the conspecific treatment (gg) to show a relationship of decreasing offspring weight with increasing bb size - which is alomst surely not reflective of reality. This is the main reason I chose to also exclude point 4 from the currently presented analysis.  

```{r}

f2.outliers <- f2.validate[-c(60,74),]
# removing points 79 and 98.

m1 <- lm(Scope.size ~ Sex + Treatment.ID + BB.weight + Treatment.ID:BB.weight + Sex:Treatment.ID, data = f2.outliers)

m2 <- lm(Scope.size ~ Sex + Treatment.ID, data = f2.outliers)

m3 <- lm(Scope.size ~ Sex + Treatment.ID + BB.weight + BB.weight:Treatment.ID, data = f2.outliers)

anova(m2,m1)
# reject reduced model m2 - close call though.

anova(m3,m1)
# Also a close call, but accept reduced model.

Anova(m3)

calc.relimp(m3, type = "lmg")
# Even when bb weight and bbweight:treatment is included, treatment effect by itself explains ~10% of variance still. About the same amount as in the reduced model with just sex and treatment (shown below).

ggplot(f2.outliers, aes(x = BB.weight, y = Scope.size, color = Treatment.ID)) + geom_point() + labs(y = "Offspring size", x = "BB weight") + geom_smooth(method = "lm")

```

> Case where observation 4 is left, plot shows an unrealistic seeming relationship for gg treatment because of this one point. 

```{r}

f2.outliers <- f2.validate[-c(4,74,60),]
# removing points 4, 79, and 98.

m1 <- lm(Scope.size ~ Sex + Treatment.ID + BB.weight + Treatment.ID:BB.weight + Sex:Treatment.ID, data = f2.outliers)

m2 <- lm(Scope.size ~ Sex + Treatment.ID, data = f2.outliers)

m3 <- lm(Scope.size ~ Sex + Treatment.ID + BB.weight + BB.weight:Treatment.ID, data = f2.outliers)

anova(m2,m1)
anova(m2,m3)
# Can't reject null hypothesis that reduced model explains the data well enough.

calc.relimp(m2, type = "lmg")

ggplot(f2.outliers, aes(x = BB.weight, y = Scope.size, color = Treatment.ID)) + geom_point() + labs(y = "Offspring size", x = "BB weight") + geom_smooth(method = "lm")

```

>  Now when observation 4 is taken out, the interaction that was significant before no longer is and the maternal effect part of the model drops out. But to me this looks like a correct decision because the plot looks much better.

```{r}

ggplot(f2.outliers, aes(x = Maternal.size, y = Scope.size)) + geom_point() + geom_jitter() + geom_smooth(method = "lm") + labs(y = "Offspring size", x = "Maternal size")

```

> But, even when all 3 of the most suspicious points are dropped, still see the trend we expect (which was also reflected in the raw data) - larger moms generally have larger offspring.